### 2020-05-22 literature.tex

\section{Misc. Literature Quotes}

In February 2014 on the Big Island, the utility reports that 10\% of its circuits hit unstable levels \cite{trabish_solar_2014}.

Photovoltaics affect power quality by introducing harmonics \cite{anurangi_effects_2017}. The power feed-in of PV generation in rural low-voltage grids can influence power quality (PQ) as well as facility operation and reliability \cite{rita_pinto_impact_2016}.

Models, not actual installations \cite{anurangi_effects_2017,bayindir_effects_2016,farhoodnea_power_2012,shafiullah_experimental_2014}. Or data taken for only one week \cite{kucuk_assessment_2013}.

``across all business sectors, the U.S. economy is losing between \$104 billion and \$164 billion a year to outages and another \$15 billion to \$24 billion to PQ phenomena'' \cite{elphick_summary_2015} India lost more than \$9.6 billion in 2008 due to power quality, and Europe lost \$150 billion per year \cite{laskar_power_2012}.

A study of the island of Porto Santo in Portugal found that intermittent nature of installed photovoltaic and wind energy could result in a potential drop in frequency of 12 Hz, lasting as long as 7 seconds \cite{delgado_solutions_2011}.

Reduced input voltage can cause excessive power supply heat dissipation, resulting in reduced mean time between failures (MTBF). In addition, rectifiers and DC-to-DC converter switching transistors draw high-peak currents, which raise their junction temperatures. These temperature excursions take a toll on semiconductor longevity. High input voltage can also puncture a power supply's rectifier and switching transistor junctions, causing MTBF reduction and eventual breakdown. High-voltage transients lasting microseconds can permanently wreck the power supply and its electronic equipment load. \cite{dedad_when_2008}.

Solar panels connected to low voltage networks will result in over-voltages, the switching frequency of the converters in wind turbines causes high-frequency signals flowing into the grid, and harmonics are generated by EV chargers. \cite{zavoda_power_2018}