### 2020-05-31 07-opq-mauka.tex
Started on mauka

\section{OPQ Mauka}
\label{sec:opq-mauka}

The previous sections discussed the design of OPQ Box, a custom hardware device for collecting four important measures of power quality, and OPQ Makai, a novel, hybrid centralized/decentralized data acquisition scheme which involves two-way communication between the OPQ Boxes.  As a result of these two systems, the OPQ sensor network services will have the ability to collect and analyze high fidelity data about power quality anomalies in a cost-effective, scalable fashion.

There are remaining challenges to creating a useful power quality sensor network. First, the data provided by OPQ Boxes is low-level, "primitive" data consisting of either features (i.e. frequency, voltage, THD, and transients) or waveform data. But what we actually want is actionable insights into grid stability. For example, we might want to know if a given anomalous data value is actually detrimental, or we might wnat to be able to predict when a power quality event might occur in the future based upon the recognition of cyclical events in the historical data.

A second challenge is data volume. Although OPQ Box and OPQ Makai provide a scalable mechanism for communicating power quality data to the cloud services, it is still the case that, over time, a substantial amount of data will accumulate. One strategy is to simply store all of the data sent to the cloud forever. This means that data storage requirements will increase monotonically over time, making the sensor network more costly to maintain the longer it is in place. An alternative strategy is to implement an algorithm to identify uninteresting (or no longer interesting) data and discard it.  Ideally, such an algorithm would enable OPQ sensor network designers to calculate an upper bound on the total amount of cloud storage required as a function of the number of nodes (OPQ Boxes) in the network.

OPQ Mauka addresses both of these issues. First, OPQ Mauka provides a multi-leveled representation for structuring and processing DSN data. The structure and processing at each level is designed with the explicit goal of turning low-level data into actionable insights. Second, each level in the framework implements a "time-to-live" (TTL) strategy for data within the level. This strategy states that data must either progress upwards through the levels towards more abstract, useful representations within a fixed time window, or be discarded and lost forever. The TTL strategy is useful because when implemented, it allows DSN designers to calculate upper bounds on data storage at each level of the framework and supports graceful degradation of DSN performance.