### 2020-07-18 cover-letter/reviewer-1.tex
revision complete.

\documentclass[12pt]{letter}
\usepackage{times}
\usepackage[left=1in,top=1in,right=1in,bottom=1in]{geometry}
\signature{Philip Johnson}
\address{Information \& Computer Sciences \\ University of Hawaii \\ Honolulu, HI USA 96822}
\longindentation=0pt
\begin{document}

\begin{letter}{}

{\bf Comments from Reviewer 1:}

{\em (1) I think this article should be shortened significantly. The first sections are very descriptive. It's more a project report than a scientific article. Only the section on the results (section 8) is more research oriented. Therefore I suggest to rework the article such that there is much more focus on the scientific aspects and results.}

Thank you for your comments.

We shortened the article from 32 pages to 24 pages.

We modified the structure to provide a more traditional scientific orientation: Introduction, Methods, Related Work, Results and Discussion, Conclusions.

We believe these changes provide more focus on scientific aspects and results.

\end{letter}
\end{document}