### 2020-07-18 cover-letter/reviewer-2.tex

\documentclass[12pt]{letter}
\usepackage{times}
\usepackage[left=1in,top=1in,right=1in,bottom=1in]{geometry}
\signature{Philip Johnson}
\address{Information \& Computer Sciences \\ University of Hawaii \\ Honolulu, HI USA 96822}
\longindentation=0pt
\begin{document}

\begin{letter}{}

{\bf Comments from Reviewer 2:}

{\em (1) In my opinion the paper is too long. Getting through the whole paper is very tiring. It seems that the maximum length of a scientific article (not including appendix) should be 20–22 pages. Maybe some of the descriptions can be moved to the appendix.}

Thank you for your comments.

We shortened the article from 32 pages to 24 pages.

{\em (2) I suggest a typical layout used in the articles: Introduction, Methods, Results and Discussion, Conclusions. This template is also recommended by Journal (see Instructions for Authors).}

We modified the structure to provide a more traditional scientific format: Introduction, Methods, Related Work, Results and Discussion, Conclusions and Future Directions.

{\em (3) Because this is a scientific article, the aims of the work and the research gap should be more clearly highlighted. I also suggest to improve the conclusions, delete the description of the work carried out (it is already well known) and focus on the most important conclusions and the future research. }

The introduction now highlights the aims of the work and the research gap (hypotheses to be tested).

Conclusion section has been improved.

{\em (4) The quality of Figures should be corrected, their descriptions, especially axis descriptions are unreadable.}

Some figures have been deleted as part of the shortening process. Those that remain have been checked for readability.

\end{letter}
\end{document}